---
title: "OU Benchmark Figures"
author: "Filippo Monti"
format:
  html:
    toc: true
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
options(stringsAsFactors = FALSE)

required_packages <- c(
  "dplyr", "forcats", "fs", "ggplot2", "patchwork", "readr",
  "scales", "showtext", "stringr", "sysfonts", "tibble", "tidyr"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages) > 0L) {
  stop(
    "Install missing packages before rendering: ",
    paste(missing_packages, collapse = ", "),
    call. = FALSE
  )
}

suppressPackageStartupMessages({
  lapply(required_packages, library, character.only = TRUE)
})

benchmark_root <- fs::path_norm(Sys.getenv(
  "OU_BENCHMARK_ROOT",
  ".."
))
if (!fs::dir_exists(fs::path(benchmark_root, "benchmarks"))) {
  stop(
    "Could not locate the public-code repo root (expected a benchmarks/ ",
    "subdir) from ", benchmark_root, ". Render this file from ",
    "public-code/figures/, or set OU_BENCHMARK_ROOT explicitly.",
    call. = FALSE
  )
}
benchmark_data_path <- fs::path(benchmark_root, "benchmarks")
branch_results_path <- fs::path(benchmark_data_path, "branch", "results")
figure_output_paths <- fs::path_norm(Sys.getenv(
  "OU_FIGURE_OUTPUT",
  fs::path(benchmark_root, "figures", "output")
))

fs::dir_create(c(
  figure_output_paths,
  branch_results_path
))

# CMU Serif / Latin Modern Math give the closest visual match to the
# manuscript's rendered figures. Both fall back gracefully (to "serif") if
# not installed locally; install CMU Serif (https://www.fontsquirrel.com/fonts/computer-modern)
# and Latin Modern Math (https://www.gust.org.pl/projects/e-foundry/lm-math)
# to reproduce the shipped PDFs byte-for-byte.
cmu_font_candidates <- list(
  regular = c(
    fs::path_home("Library", "Fonts", "cmunrm.otf"),
    "/usr/share/fonts/opentype/cmu/cmunrm.otf",
    "/usr/share/fonts/truetype/cmu/cmunrm.otf"
  ),
  italic = c(
    fs::path_home("Library", "Fonts", "cmunti.otf"),
    "/usr/share/fonts/opentype/cmu/cmunti.otf",
    "/usr/share/fonts/truetype/cmu/cmunti.otf"
  ),
  bold = c(
    fs::path_home("Library", "Fonts", "cmunbx.otf"),
    "/usr/share/fonts/opentype/cmu/cmunbx.otf",
    "/usr/share/fonts/truetype/cmu/cmunbx.otf"
  ),
  bolditalic = c(
    fs::path_home("Library", "Fonts", "cmunbi.otf"),
    "/usr/share/fonts/opentype/cmu/cmunbi.otf",
    "/usr/share/fonts/truetype/cmu/cmunbi.otf"
  )
)
find_first_font <- function(paths) {
  existing <- paths[fs::file_exists(paths)]
  if (length(existing) > 0) existing[1] else NA_character_
}
cmu_paths <- lapply(cmu_font_candidates, find_first_font)
if (!is.na(cmu_paths$regular)) {
  sysfonts::font_add(
    "CMU Serif",
    regular = cmu_paths$regular,
    italic = cmu_paths$italic,
    bold = cmu_paths$bold,
    bolditalic = cmu_paths$bolditalic
  )
}

lm_math_candidates <- c(
  "/Library/TeX/Root/texmf-dist/fonts/opentype/public/lm-math/latinmodern-math.otf",
  "/usr/share/fonts/opentype/public/lm-math/latinmodern-math.otf"
)
lm_math_path <- find_first_font(lm_math_candidates)
if (!is.na(lm_math_path)) {
  sysfonts::font_add("lmmath", lm_math_path)
}

stix_text_regular_candidates <- c(
  "/System/Library/Fonts/Supplemental/STIXTwoText.ttf",
  "/usr/share/fonts/opentype/stix/STIXTwoText-Regular.otf"
)
stix_text_italic_candidates <- c(
  "/System/Library/Fonts/Supplemental/STIXTwoText-Italic.ttf",
  "/usr/share/fonts/opentype/stix/STIXTwoText-Italic.otf"
)
stix_text_regular <- stix_text_regular_candidates[fs::file_exists(stix_text_regular_candidates)][1]
stix_text_italic <- stix_text_italic_candidates[fs::file_exists(stix_text_italic_candidates)][1]
if (!is.na(stix_text_regular) && !is.na(stix_text_italic)) {
  sysfonts::font_add(
    "STIXTwoText",
    regular = stix_text_regular,
    italic = stix_text_italic
  )
}

stix_math_candidates <- c(
  "/System/Library/Fonts/Supplemental/STIXTwoMath.otf",
  "/usr/share/fonts/opentype/stix/STIXTwoMath-Regular.otf"
)
stix_math_path <- stix_math_candidates[fs::file_exists(stix_math_candidates)][1]
if (!is.na(stix_math_path)) {
  sysfonts::font_add("STIXTwoMath", regular = stix_math_path)
}
math_annotation_family <- if ("STIXTwoText" %in% sysfonts::font_families()) {
  "STIXTwoText"
} else if ("STIXTwoMath" %in% sysfonts::font_families()) {
  "STIXTwoMath"
} else {
  "serif"
}
showtext::showtext_auto(enable = TRUE)

read_benchmark_csv <- function(...) {
  readr::read_csv(fs::path(...), show_col_types = FALSE)
}

save_benchmark_plot <- function(plot, filename, width, height) {
  paths <- fs::path(figure_output_paths, filename)
  for (path in paths) {
    ggplot2::ggsave(path, plot = plot, width = width, height = height, dpi = 300)
  }
  invisible(paths)
}

save_benchmark_pdf <- function(plot, basename, width, height) {
  save_benchmark_plot(plot, paste0(basename, ".pdf"), width, height)
}

save_benchmark_text <- function(lines, filename) {
  paths <- fs::path(figure_output_paths, filename)
  for (path in paths) {
    writeLines(lines, path)
  }
  invisible(paths)
}

paper_theme <- function(base_size = 9) {
  theme_classic(base_size = base_size, base_family = "serif") +
    theme(
      axis.line = element_line(linewidth = 0.3),
      axis.ticks = element_line(linewidth = 0.3),
      axis.text = element_text(color = "black"),
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.35),
      strip.background = element_rect(fill = "grey93", color = "black", linewidth = 0.3),
      strip.text = element_text(face = "bold"),
      plot.title = element_text(face = "bold", hjust = 0),
      plot.subtitle = element_text(color = "grey25")
    )
}

kernel_colors <- c(
  "SSBP (R invertible, LDU)" = "#2F67B1",
  "SSBP (R orthogonal)" = "#228B22",
  "SSBP block only" = "#6F7B82",
  "Dense" = "#D4960A",
  "Known Schur" = "#9B4356",
  "Schur Parlett" = "#D67C2F"
)

kernel_alpha <- c(
  "SSBP (R invertible, LDU)" = 1,
  "SSBP (R orthogonal)" = 1,
  "SSBP block only" = 0.7,
  "Dense" = 1,
  "Known Schur" = 1,
  "Schur Parlett" = 1
)

kernel_linetypes <- c(
  "SSBP (R invertible, LDU)" = "solid",
  "SSBP (R orthogonal)" = "42",
  "SSBP block only" = "solid",
  "Dense" = "longdash",
  "Known Schur" = "longdash",
  "Schur Parlett" = "longdash"
)
```

# Julia Kernel Benchmarks

The kernel panels read the precomputed CSV files in `benchmarks/branch/results`. The current
figure filters those files to `K = 4, 8, 16, 32, 64`; no Julia benchmarks are run
by this document.

```{r kernel-data}
schur_smbp <- read_benchmark_csv(branch_results_path, "schur_vs_smbp_exp_adjoint_results.csv")
edge_push <- read_benchmark_csv(branch_results_path, "edge_message_push_results.csv")
lyapunov <- read_benchmark_csv(branch_results_path, "lyapunov_stationary_variance_results.csv")
branch_lyapunov <- read_benchmark_csv(branch_results_path, "branch_likelihood_with_lyapunov_results.csv")
branch_cov_adjoint <- read_benchmark_csv(branch_results_path, "branch_covariance_adjoint_results.csv")

keep_k <- c(4L, 8L, 16L, 32L, 64L)

ensure_benchmark_column <- function(data, column, fallback) {
  if (!column %in% names(data)) {
    data[[column]] <- data[[fallback]]
  }
  data
}

schur_smbp <- schur_smbp %>%
  ensure_benchmark_column("smbp_exp_ldu_with_rotation_ms", "smbp_exp_with_rotation_ms") %>%
  ensure_benchmark_column("smbp_adj_ldu_with_rotation_ms", "smbp_adj_with_rotation_ms")
lyapunov <- lyapunov %>%
  ensure_benchmark_column("smbp_ldu_lyap_ms", "smbp_rotation_lyap_ms")
edge_push <- edge_push %>%
  ensure_benchmark_column("smbp_ldu_full_edge_with_rotation_ms", "smbp_full_edge_with_rotation_ms")
branch_lyapunov <- branch_lyapunov %>%
  ensure_benchmark_column("smbp_branch_lyap_ldu_rotation_ms", "smbp_branch_lyap_rotation_ms")

kernel_timings <- bind_rows(
  schur_smbp %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Matrix exponential",
      Dense = dense_exp_ms,
      `Known Schur` = schur_exp_ms,
      `Schur Parlett` = schur_parlett_exp_ms,
      `SSBP block only` = smbp_block_exp_ms,
      `SSBP (R orthogonal)` = smbp_exp_with_rotation_ms,
      `SSBP (R invertible, LDU)` = smbp_exp_ldu_with_rotation_ms
    ),
  schur_smbp %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Matrix exp. adjoint",
      Dense = dense_adj_ms,
      `Known Schur` = schur_adj_ms,
      `SSBP block only` = smbp_adj_apply_ms,
      `SSBP (R orthogonal)` = smbp_adj_with_rotation_ms,
      `SSBP (R invertible, LDU)` = smbp_adj_ldu_with_rotation_ms
    ),
  lyapunov %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Lyapunov Equation",
      Dense = dense_lyap_ms,
      `Known Schur` = schur_known_lyap_ms,
      `SSBP (R orthogonal)` = smbp_rotation_lyap_ms,
      `SSBP (R invertible, LDU)` = smbp_ldu_lyap_ms
    ),
  edge_push %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Full edge message",
      Dense = dense_full_edge_ms,
      `Known Schur` = schur_known_basis_full_edge_ms,
      `SSBP block only` = smbp_block_full_edge_ms,
      `SSBP (R orthogonal)` = smbp_full_edge_with_rotation_ms,
      `SSBP (R invertible, LDU)` = smbp_ldu_full_edge_with_rotation_ms
    ),
  branch_lyapunov %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Branch + Lyapunov",
      Dense = dense_branch_lyap_ms,
      `Known Schur` = schur_branch_lyap_ms,
      `SSBP (R orthogonal)` = smbp_branch_lyap_rotation_ms,
      `SSBP (R invertible, LDU)` = smbp_branch_lyap_ldu_rotation_ms
    )
) %>%
  pivot_longer(
    cols = -c(K, benchmark),
    names_to = "method",
    values_to = "milliseconds",
    values_drop_na = TRUE
  ) %>%
  filter(milliseconds > 0) %>%
  mutate(
    method = factor(method, levels = names(kernel_colors)),
    benchmark = factor(
      benchmark,
      levels = c(
        "Matrix exponential", "Matrix exp. adjoint",
        "Lyapunov Equation", "Full edge message",
        "Branch + Lyapunov"
      )
    )
  )
```

```{r kernel-figure, fig.width=10, fig.height=6.8}
fnt <- "CMU Serif"
fsz <- 10
fszT <- 12

kernel_methods <- names(kernel_colors)
kernel_lw <- c(
  "SSBP (R invertible, LDU)" = 1.0,
  "SSBP (R orthogonal)" = 0.95,
  "SSBP block only" = 0.8,
  "Dense" = 0.5,
  "Known Schur" = 0.5,
  "Schur Parlett" = 0.5
)
kernel_lty_legend <- c(
  "SSBP (R invertible, LDU)" = "solid",
  "SSBP (R orthogonal)" = "42",
  "SSBP block only" = "solid",
  "Dense" = "32",
  "Known Schur" = "32",
  "Schur Parlett" = "32"
)

kernel_panel_limits <- function(data, pad = 1.8) {
  values <- data$milliseconds
  values <- values[is.finite(values) & values > 0]
  if (length(values) == 0L) {
    return(c(1e-4, 1))
  }
  c(min(values) / pad, max(values) * pad)
}

kernel_curve_y_at <- function(data, method_value, x_value) {
  curve <- data %>%
    filter(method == method_value, is.finite(milliseconds), milliseconds > 0) %>%
    arrange(K)
  if (nrow(curve) == 0L) {
    return(NA_real_)
  }
  10^approx(
    x = log2(curve$K),
    y = log10(curve$milliseconds),
    xout = log2(x_value),
    rule = 2
  )$y
}

kernel_cubic_reference <- function(data, anchor_method,
                                   anchor_k = 64L, start_k = 16L) {
  anchor <- data %>%
    filter(K == anchor_k, method == anchor_method, milliseconds > 0) %>%
    slice(1L)
  if (nrow(anchor) == 0L) {
    return(tibble())
  }
  k_values <- keep_k[keep_k >= start_k & keep_k <= anchor_k]
  tibble(
    anchor_method = anchor_method,
    K = k_values,
    milliseconds = anchor$milliseconds[[1L]] * (k_values / anchor_k)^3
  )
}

kernel_cubic_reference_label <- function(data, anchor_method,
                                         anchor_k = 64L, label_k = 20,
                                         label_scale = 1.35) {
  anchor <- data %>%
    filter(K == anchor_k, method == anchor_method, milliseconds > 0) %>%
    slice(1L)
  if (nrow(anchor) == 0L) {
    return(tibble())
  }
  tibble(
    anchor_method = anchor_method,
    K = label_k,
    milliseconds = label_scale * anchor$milliseconds[[1L]] * (label_k / anchor_k)^3,
    label = "K^3"
  )
}

kernel_best_arrow_label <- function(data, k_value, y_start, y_end) {
  y_low <- min(y_start, y_end)
  y_high <- max(y_start, y_end)
  y_mid <- sqrt(y_start * y_end)
  y_limits <- kernel_panel_limits(data)

  candidates <- tibble(
    x_label = c(k_value * 1.14, k_value / 1.14, k_value * 1.08, k_value / 1.08),
    y_label = c(y_mid, y_mid, y_low / 1.55, y_low / 1.55),
    hjust = c(0, 1, 0, 1),
    vjust = c(0.5, 0.5, 1, 1),
    placement = c("right", "left", "below-right", "below-left")
  ) %>%
    filter(
      x_label >= min(keep_k) * 0.92,
      x_label <= max(keep_k) * 1.10,
      y_label >= y_limits[[1L]],
      y_label <= y_limits[[2L]]
    )

  if (nrow(candidates) == 0L) {
    return(tibble(x_label = k_value, y_label = y_mid, hjust = 0.5, vjust = 0.5))
  }

  methods <- unique(as.character(data$method))
  candidates %>%
    rowwise() %>%
    mutate(
      clearance = min(abs(log10(y_label) - log10(vapply(
        methods,
        function(method_value) kernel_curve_y_at(data, method_value, x_label),
        numeric(1)
      ))), na.rm = TRUE),
      score = clearance + if_else(str_detect(placement, "below"), 0.04, 0)
    ) %>%
    ungroup() %>%
    arrange(desc(score)) %>%
    slice(1L) %>%
    select(x_label, y_label, hjust, vjust)
}

kernel_panel <- function(data, title) {
  d <- data %>%
    filter(method %in% kernel_methods) %>%
    mutate(method = factor(method, levels = kernel_methods))
  cubic_reference <- bind_rows(
    kernel_cubic_reference(d, "SSBP (R invertible, LDU)"),
    kernel_cubic_reference(d, "Known Schur")
  )
  cubic_reference_label <- bind_rows(
    kernel_cubic_reference_label(d, "SSBP (R invertible, LDU)", label_k = 20),
    kernel_cubic_reference_label(d, "Known Schur", label_k = 18)
  )

  p <- ggplot(
    d,
    aes(
      K,
      milliseconds,
      colour = method,
      shape = method,
      alpha = method,
      linetype = method,
      linewidth = method
    )
  ) +
    geom_line(
      data = cubic_reference,
      aes(x = K, y = milliseconds, group = anchor_method),
      inherit.aes = FALSE,
      colour = "grey35",
      linetype = "22",
      linewidth = 0.35
    ) +
    geom_line() +
    geom_point(
      data = d,
      aes(x = K, y = milliseconds),
      inherit.aes = FALSE,
      colour = "white",
      alpha = 1,
      size = 1.9,
      stroke = 0
    ) +
    geom_point(size = 1.9) +
    scale_x_continuous(
      trans = "log2",
      breaks = keep_k,
      labels = as.character(keep_k),
      expand = expansion(mult = c(0.05, 0.14))
    ) +
    scale_y_log10(
      breaks = 10^(-6:6),
      labels = function(x) {
        pow <- log10(x)
        ifelse(
          pow %% 2 == 0,
          scales::math_format(10^.x)(pow),
          ""
        )
      },
      minor_breaks = NULL
    ) +
    coord_cartesian(ylim = kernel_panel_limits(d)) +
    scale_colour_manual(values = kernel_colors[kernel_methods], name = NULL) +
    scale_shape_manual(values = rep(16L, length(kernel_methods)), name = NULL) +
    scale_alpha_manual(values = kernel_alpha[kernel_methods], name = NULL) +
    scale_linetype_manual(values = kernel_linetypes[kernel_methods], name = NULL) +
    scale_linewidth_manual(values = kernel_lw[kernel_methods], name = NULL) +
    labs(x = NULL, y = NULL, title = title) +
    theme_classic(base_size = fszT, base_family = fnt) +
    theme(
      aspect.ratio = 1,
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
      panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_blank(),
      axis.ticks = element_line(linewidth = 0.3),
      axis.text = element_text(size = fsz, family = fnt, color = "black"),
      axis.title = element_text(size = fszT, family = fnt, color = "black"),
      plot.title = element_text(
        hjust = 0.5,
        size = fszT,
        family = fnt,
        face = "bold",
        margin = margin(0, 0, 1, 0, "pt")
      ),
      plot.margin = margin(-2, 4, -2, 1, "pt"),
      legend.position = "none"
    )

  if (nrow(cubic_reference_label) > 0L) {
    p <- p +
      geom_text(
        data = cubic_reference_label,
        aes(x = K, y = milliseconds, label = label),
        inherit.aes = FALSE,
        parse = TRUE,
        size = 3.0,
        family = fnt,
        colour = "grey25"
      )
  }

  arrow_rows <- bind_rows(lapply(c(8L, 32L), function(k_value) {
    smbp_row <- d %>%
      filter(K == k_value, method == "SSBP (R invertible, LDU)")
    competitor_row <- d %>%
      filter(K == k_value, !str_detect(as.character(method), "^SSBP")) %>%
      arrange(milliseconds) %>%
      slice(1L)

    if (nrow(smbp_row) == 0L || nrow(competitor_row) == 0L) {
      return(tibble())
    }

    competitor_ms <- competitor_row$milliseconds[[1L]]
    smbp_ms <- smbp_row$milliseconds[[1L]]
    label_position <- kernel_best_arrow_label(d, k_value, competitor_ms, smbp_ms)
    arrow_gap <- 1.28
    tibble(
      K = k_value,
      y_start = competitor_ms / arrow_gap,
      y_end = smbp_ms * arrow_gap,
      x_label = label_position$x_label,
      y_label = label_position$y_label,
      hjust = label_position$hjust,
      vjust = label_position$vjust,
      label = if (competitor_ms / smbp_ms >= 10) {
        paste0(round(competitor_ms / smbp_ms), "x")
      } else {
        paste0(signif(competitor_ms / smbp_ms, 2), "x")
      }
    )
  }))

  if (nrow(arrow_rows) > 0L) {
    p <- p +
      geom_segment(
        data = arrow_rows,
        aes(x = K, xend = K, y = y_start, yend = y_end),
        inherit.aes = FALSE,
        colour = "black",
        linewidth = 0.3,
        arrow = arrow(type = "open", length = grid::unit(0.05, "inches"))
      ) +
      geom_text(
        data = arrow_rows,
        aes(x = x_label, y = y_label, label = label, hjust = hjust, vjust = vjust),
        inherit.aes = FALSE,
        size = 3.2,
        family = fnt,
        colour = "black"
      )
  }

  p
}

kernel_y_lab <- wrap_elements(
  grid::textGrob(
    "Time (ms)",
    rot = 90,
    gp = grid::gpar(fontsize = fszT, fontfamily = fnt)
  )
)

kernel_x_lab <- wrap_elements(
  full = grid::textGrob(
    "State dimension (K)",
    y = grid::unit(1, "npc"),
    just = c("centre", "top"),
    gp = grid::gpar(fontsize = fszT, fontfamily = fnt)
  )
)

kernel_leg_panel <- function(methods = kernel_methods) {
  c1x <- 0.10
  c2x <- 0.43
  seg <- 0.030
  gap <- 0.018
  proposed <- methods[str_detect(methods, "^SSBP")]
  benchmarks <- setdiff(methods, proposed)

  leg_df <- tibble(
    method = factor(c(proposed, benchmarks), levels = names(kernel_colors)),
    label = c(proposed, benchmarks),
    x_seg0 = c(rep(c1x, length(proposed)), rep(c2x, length(benchmarks))),
    y = c(
      seq(1.6, by = -0.8, length.out = length(proposed)),
      seq(1.9, by = -0.8, length.out = length(benchmarks))
    )
  ) %>%
    mutate(x_lab = x_seg0 + seg + gap)
  leg_titles <- tibble(
    label = c("Proposed methods", "Dense baselines"),
    x = c(c1x, c2x),
    y = c(2.9, 2.9)
  )

  ggplot() +
    geom_text(
      data = leg_titles,
      aes(x = x, y = y, label = label),
      hjust = 0,
      size = 3.5,
      family = fnt,
      fontface = "bold",
      colour = "grey10",
      inherit.aes = FALSE
    ) +
    geom_segment(
      data = leg_df,
      aes(
        x = x_seg0,
        xend = x_seg0 + seg,
        y = y,
        yend = y,
        colour = method,
        alpha = method,
        linetype = method,
        linewidth = method
      ),
      show.legend = FALSE,
      inherit.aes = FALSE
    ) +
    geom_text(
      data = leg_df,
      aes(x = x_lab, y = y, label = label),
      hjust = 0,
      size = 3.5,
      family = fnt,
      colour = "grey20",
      inherit.aes = FALSE
    ) +
    scale_colour_manual(values = kernel_colors[methods]) +
    scale_alpha_manual(values = kernel_alpha[methods]) +
    scale_linetype_manual(values = kernel_lty_legend[methods]) +
    scale_linewidth_manual(values = kernel_lw[methods]) +
    scale_x_continuous(limits = c(0, 0.95)) +
    scale_y_continuous(limits = c(-0.7, 3.3)) +
    theme_void() +
    theme(legend.position = "none")
}

kernel_titles <- levels(kernel_timings$benchmark)
kernel_panels <- lapply(kernel_titles, function(title) {
  kernel_panel(filter(kernel_timings, benchmark == title), title)
})

kernel_fig <- wrap_plots(
  kernel_y_lab,
  kernel_panels[[1]],
  kernel_panels[[2]],
  kernel_panels[[3]],
  kernel_panels[[4]],
  kernel_panels[[5]],
  kernel_x_lab,
  kernel_leg_panel(),
  design = "
ABCD
#EF#
#GG#
#HH#
",
  widths = c(0.06, 1, 1, 1),
  heights = c(1, 1, 0.05, 0.26)
)

kernel_fig
```

```{r exp-adjoint-cached-figure, fig.width=10, fig.height=5.8}
schur_smbp_cached <- schur_smbp %>%
  ensure_benchmark_column("schur_basis_exp_ms", "schur_exp_ms") %>%
  ensure_benchmark_column("schur_basis_adj_ms", "schur_adj_ms")

lyapunov_cached <- lyapunov %>%
  ensure_benchmark_column("schur_basis_lyap_ms", "schur_known_lyap_ms") %>%
  ensure_benchmark_column("smbp_basis_lyap_ms", "smbp_rotation_lyap_ms") %>%
  ensure_benchmark_column("smbp_ldu_basis_lyap_ms", "smbp_ldu_lyap_ms")

focused_methods <- c(
  "H-SSBP",
  # "H-SSBP (R orthogonal)",
  "Known Schur",
  "Dense (cached)"
)

focused_colors <- c(
  "H-SSBP" = kernel_colors[["SSBP (R invertible, LDU)"]],
  # "H-SSBP (R orthogonal)" = "#86A9CF",
  "Known Schur" = kernel_colors[["Known Schur"]],
  "Dense (cached)" = kernel_colors[["Dense"]]
)

focused_linetypes <- c(
  "H-SSBP" = "solid",
  # "H-SSBP (R orthogonal)" = "solid",
  "Known Schur" = kernel_lty_legend[["Known Schur"]],
  "Dense (cached)" = kernel_lty_legend[["Dense"]]
)

focused_lw <- c(
  "H-SSBP" = 1.0,
  # "H-SSBP (R orthogonal)" = 0.55,
  "Known Schur" = 0.65,
  "Dense (cached)" = 0.65
)

exp_adjoint_cached <- bind_rows(
  schur_smbp_cached %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Matrix exponential",
      `H-SSBP` = smbp_block_exp_ms,
      # `H-SSBP (R orthogonal)` = smbp_block_exp_ms,
      `Known Schur` = schur_basis_exp_ms,
      `Dense (cached)` = dense_exp_ms
    ),
  schur_smbp_cached %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Matrix exp. adjoint",
      `H-SSBP` = smbp_adj_apply_ms,
      # `H-SSBP (R orthogonal)` = smbp_adj_apply_ms,
      `Known Schur` = schur_basis_adj_ms,
      `Dense (cached)` = dense_adj_ms
    ),
  lyapunov_cached %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Lyapunov equation",
      `H-SSBP` = smbp_ldu_basis_lyap_ms,
      # `H-SSBP (R orthogonal)` = smbp_basis_lyap_ms,
      `Known Schur` = schur_basis_lyap_ms,
      `Dense (cached)` = dense_lyap_ms
    ),
  edge_push %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Full edge message",
      `H-SSBP` = smbp_ldu_full_edge_with_rotation_ms,
      # `H-SSBP (R orthogonal)` = smbp_full_edge_with_rotation_ms,
      `Known Schur` = schur_known_basis_full_edge_ms,
      `Dense (cached)` = dense_full_edge_ms
    ),
  branch_cov_adjoint %>%
    filter(K %in% keep_k) %>%
    transmute(
      K,
      benchmark = "Full edge message gradient",
      `H-SSBP` = smbp_basis_branch_cov_adj_ms,
      `Known Schur` = schur_basis_branch_cov_adj_ms,
      `Dense (cached)` = dense_branch_cov_adj_ms
    )
) %>%
  pivot_longer(
    cols = -c(K, benchmark),
    names_to = "method",
    values_to = "milliseconds",
    values_drop_na = TRUE
  ) %>%
  filter(is.finite(milliseconds), milliseconds > 0) %>%
  mutate(
    method = factor(method, levels = focused_methods),
    benchmark = factor(
      benchmark,
      levels = c(
        "Matrix exponential", "Matrix exp. adjoint",
        "Lyapunov equation", "Full edge message", "Full edge message gradient"
      )
    )
  )

focused_panel_limits <- function(data, pad = 1.7) {
  values <- data$milliseconds
  values <- values[is.finite(values) & values > 0]
  c(min(values) / pad, max(values) * pad)
}

focused_y_labels <- function(x) {
  pow <- log10(x)
  ifelse(
    pow %% 2 == 0,
    scales::math_format(10^.x)(pow),
    ""
  )
}

focused_best_arrow_label <- function(data, k_value, y_start, y_end, y_limits = NULL) {
  y_low <- min(y_start, y_end)
  y_mid <- sqrt(y_start * y_end)
  if (is.null(y_limits)) {
    y_limits <- focused_panel_limits(data)
  }

  candidates <- tibble(
    x_label = c(k_value * 1.14, k_value / 1.14, k_value * 1.08, k_value / 1.08),
    y_label = c(y_mid, y_mid, y_low / 1.55, y_low / 1.55),
    hjust = c(0, 1, 0, 1),
    vjust = c(0.5, 0.5, 1, 1),
    placement = c("right", "left", "below-right", "below-left")
  ) %>%
    filter(
      x_label >= min(keep_k) * 0.92,
      x_label <= max(keep_k) * 1.10,
      y_label >= y_limits[[1L]],
      y_label <= y_limits[[2L]]
    )

  if (nrow(candidates) == 0L) {
    return(tibble(x_label = k_value, y_label = y_mid, hjust = 0.5, vjust = 0.5))
  }

  methods <- unique(as.character(data$method))
  candidates %>%
    rowwise() %>%
    mutate(
      clearance = min(abs(log10(y_label) - log10(vapply(
        methods,
        function(method_value) kernel_curve_y_at(data, method_value, x_label),
        numeric(1)
      ))), na.rm = TRUE),
      score = clearance + if_else(str_detect(placement, "below"), 0.04, 0)
    ) %>%
    ungroup() %>%
    arrange(desc(score)) %>%
    slice(1L) %>%
    select(x_label, y_label, hjust, vjust)
}

focused_arrow_rows <- function(data, y_limits = NULL) {
  bind_rows(lapply(c(8L, 32L), function(k_value) {
    smbp_row <- data %>%
      filter(K == k_value, method == "H-SSBP")
    competitor_row <- data %>%
      filter(K == k_value, method %in% c("Known Schur", "Dense (cached)")) %>%
      arrange(milliseconds) %>%
      slice(1L)

    if (nrow(smbp_row) == 0L || nrow(competitor_row) == 0L) {
      return(tibble())
    }

    competitor_ms <- competitor_row$milliseconds[[1L]]
    smbp_ms <- smbp_row$milliseconds[[1L]]
    is_lyapunov_panel <- unique(as.character(data$benchmark))[[1L]] == "Lyapunov equation"
    label_position <- if (is_lyapunov_panel) {
      tibble(
        x_label = k_value * 1.08,
        y_label = smbp_ms / 1.55,
        hjust = 0,
        vjust = 1
      )
    } else {
      focused_best_arrow_label(data, k_value, competitor_ms, smbp_ms, y_limits)
    }
    arrow_gap <- 1.28
    speedup <- competitor_ms / smbp_ms
    tibble(
      K = k_value,
      y_start = competitor_ms / arrow_gap,
      y_end = smbp_ms * arrow_gap,
      x_label = label_position$x_label,
      y_label = label_position$y_label,
      hjust = label_position$hjust,
      vjust = label_position$vjust,
      label = if (speedup < 5) {
        paste0(round(speedup), "x")
      } else {
        paste0(round(speedup / 5) * 5, "x")
      }
    )
  }))
}

first_three_limits <- exp_adjoint_cached %>%
  filter(benchmark %in% c("Matrix exponential", "Matrix exp. adjoint", "Lyapunov equation")) %>%
  focused_panel_limits()

# Scale text elements so they land at the same printed sizes as the
# ssbp-vs-lyapunov-param-stability figure's combo_panel_theme, once this
# 10in-wide figure is placed at 0.98\textwidth in the same 6.5in-textwidth,
# 12pt document (printed pt = raw ggplot pt * focused_scale_down). See the
# combo_scale_down/combo_font_scale comment above the
# ssbp-vs-lyapunov-param-stability-figure chunk for the derivation; this is
# the same formula evaluated at this figure's own native width (10in vs
# 9.2in) so the two figures' printed font sizes match despite differing.
focused_scale_down <- 0.98 * 6.5 / 10
focused_font_scale <- (12 / focused_scale_down) / 9
focused_title_pt <- 8 * focused_font_scale
# Ratio applied to this figure's other hand-tuned annotation/legend text
# sizes (originally tuned against base_size = 9) so they scale up by the
# same factor as the axis/title text above, instead of staying fixed while
# everything else grows.
focused_text_scale <- focused_title_pt / 9
focused_plot_title_pt <- focused_title_pt - 1 / focused_scale_down
focused_speedup_size <- 3.2 * focused_text_scale - 1 / (ggplot2::.pt * focused_scale_down)
focused_math_label_size <- focused_speedup_size - 3 / (ggplot2::.pt * focused_scale_down)

focused_panel <- function(data, title, y_limits = NULL, show_y = TRUE) {
  p <- ggplot(
    data,
    aes(
      K,
      milliseconds,
      colour = method,
      linetype = method,
      linewidth = method,
      group = method
    )
  ) +
    geom_line() +
    geom_point(size = 1.55, stroke = 0) +
    scale_x_continuous(
      trans = "log2",
      breaks = keep_k,
      labels = as.character(keep_k),
      expand = expansion(mult = c(0.05, 0.10))
    ) +
    scale_y_log10(
      breaks = 10^(-6:6),
      labels = focused_y_labels,
      minor_breaks = NULL
    ) +
    scale_colour_manual(values = focused_colors, name = NULL) +
    scale_linetype_manual(values = focused_linetypes, name = NULL) +
    scale_linewidth_manual(values = focused_lw, name = NULL) +
    labs(x = NULL, y = NULL, title = title) +
    paper_theme(base_size = focused_title_pt) +
    theme(
      legend.position = "none",
      panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(hjust = 0.5, size = focused_plot_title_pt, face = "bold"),
      axis.title = element_text(size = focused_title_pt),
      axis.text.y = if (show_y) element_text(color = "black") else element_blank(),
      axis.ticks.y = if (show_y) element_line(linewidth = 0.3) else element_blank(),
      plot.margin = margin(2, 4, 2, 4, "pt")
    )

  arrow_rows <- focused_arrow_rows(data, y_limits)
  if (nrow(arrow_rows) > 0L) {
    p <- p +
      geom_segment(
        data = arrow_rows,
        aes(x = K, xend = K, y = y_start, yend = y_end),
        inherit.aes = FALSE,
        colour = "black",
        linewidth = 0.3,
        arrow = grid::arrow(type = "open", length = grid::unit(0.05, "inches"))
      ) +
      geom_text(
        data = arrow_rows,
        aes(x = x_label, y = y_label, label = label, hjust = hjust, vjust = vjust),
        inherit.aes = FALSE,
        size = focused_speedup_size,
        family = fnt,
        colour = "black"
      )
  }

  if (identical(title, "Matrix exponential")) {
    exp_labels <- bind_rows(
      data %>%
        filter(K == 16, method == "H-SSBP") %>%
        slice(1) %>%
        transmute(K = K * 1.08, milliseconds = milliseconds * 2.6, method, label = "italic(e)^{italic(t)*italic(B)}"),
      data %>%
        filter(K == 16, method == "Known Schur") %>%
        slice(1) %>%
        transmute(K = K * 1.08, milliseconds = milliseconds * 0.48, method, label = "italic(e)^{italic(t)*italic(T)}"),
      data %>%
        filter(K == 16, method == "Dense (cached)") %>%
        slice(1) %>%
        transmute(K = K * 1.08, milliseconds = milliseconds * 2.8, method, label = "italic(e)^{italic(t)*italic(S)}")
    )

    p <- p +
      geom_text(
        data = exp_labels,
        aes(K, milliseconds, label = label, colour = method),
        inherit.aes = FALSE,
        parse = TRUE,
        show.legend = FALSE,
        size = focused_math_label_size
      )
  }

  if (identical(title, "Matrix exp. adjoint")) {
    adjoint_labels <- bind_rows(
      data %>%
        filter(K == 16, method == "H-SSBP") %>%
        slice(1) %>%
        transmute(
          K = K * 1.68,
          milliseconds = milliseconds * 0.48,
          method,
          label = "italic(F)[italic(B)]^\"*\"~italic(\"(\")*italic(R)^{italic(T)}~italic(X)~italic(R)^{-italic(T)}*italic(\")\")"
        ),
      data %>%
        filter(K == 16, method == "Known Schur") %>%
        slice(1) %>%
        transmute(
          K = K * 1.15,
          milliseconds = milliseconds * 0.18,
          method,
          label = "italic(F)[italic(T)]^\"*\"~italic(\"(\")*italic(R)^{italic(T)}~italic(X)~italic(R)^{-italic(T)}*italic(\")\")"
        ),
      data %>%
        filter(K == 16, method == "Dense (cached)") %>%
        slice(1) %>%
        transmute(
          K = K * 1.02,
          milliseconds = milliseconds * 4.2,
          method,
          label = "italic(F)[italic(S)]^\"*\"~italic(\"(\")*italic(X)*italic(\")\")"
        )
    )

    p <- p +
      geom_text(
        data = adjoint_labels,
        aes(K, milliseconds, label = label, colour = method),
        inherit.aes = FALSE,
        parse = TRUE,
        show.legend = FALSE,
        family = math_annotation_family,
        size = focused_math_label_size
      )
  }

  if (!is.null(y_limits)) {
    p <- p + coord_cartesian(ylim = y_limits)
  }
  p
}

second_row_limits <- exp_adjoint_cached %>%
  filter(benchmark %in% c("Full edge message", "Full edge message gradient")) %>%
  focused_panel_limits()

focused_leg_panel <- function(methods = focused_methods, size_up = 1.25) {
  x0 <- 0.06 * size_up
  seg <- 0.16 * size_up
  gap <- 0.05 * size_up
  entry_gap <- 0.24
  n <- length(methods)
  block_width <- 1.7 * size_up
  pad_x <- 0.3
  pad_y <- 0.75
  legend_text_size <- 3.5 * size_up * focused_text_scale - 3 / (ggplot2::.pt * focused_scale_down)

  leg_df <- tibble(
    method = factor(methods, levels = methods),
    label = methods,
    y = seq(entry_gap * (n - 1) / 2, by = -entry_gap, length.out = n)
  ) %>%
    mutate(x_lab = x0 + seg + gap)

  ggplot() +
    geom_segment(
      data = leg_df,
      aes(
        x = x0,
        xend = x0 + seg,
        y = y,
        yend = y,
        colour = method,
        linetype = method,
        linewidth = method
      ),
      show.legend = FALSE,
      inherit.aes = FALSE
    ) +
    geom_point(
      data = leg_df,
      aes(x = x0 + seg / 2, y = y, colour = method),
      size = 1.55 * size_up,
      stroke = 0,
      show.legend = FALSE,
      inherit.aes = FALSE
    ) +
    geom_text(
      data = leg_df,
      aes(x = x_lab, y = y, label = label),
      hjust = 0,
      size = legend_text_size,
      family = fnt,
      colour = "grey20",
      inherit.aes = FALSE
    ) +
    scale_colour_manual(values = focused_colors[methods]) +
    scale_linetype_manual(values = focused_linetypes[methods]) +
    scale_linewidth_manual(values = focused_lw[methods] * size_up) +
    scale_x_continuous(limits = c(-pad_x, block_width + pad_x)) +
    scale_y_continuous(limits = c(-entry_gap * (n - 1) / 2 - pad_y, entry_gap * (n - 1) / 2 + pad_y)) +
    theme_void() +
    theme(legend.position = "none")
}

shared_time_label_panel <- function(label = "Time (ms)") {
  ggplot() +
    annotate(
      "text", x = 0, y = 0, label = label,
      angle = 90, hjust = 0.5, vjust = 0.5,
      size = 3.2 * focused_text_scale, family = fnt, colour = "black"
    ) +
    scale_x_continuous(limits = c(-1, 1)) +
    scale_y_continuous(limits = c(-1, 1)) +
    theme_void()
}

panel_data <- split(exp_adjoint_cached, exp_adjoint_cached$benchmark)
exp_adjoint_cached_fig <- wrap_plots(
  shared_time_label_panel(),
  focused_panel(panel_data[["Matrix exponential"]], "Matrix exponential", first_three_limits, TRUE),
  focused_panel(panel_data[["Matrix exp. adjoint"]], "Matrix exp. adjoint", first_three_limits, FALSE),
  focused_panel(panel_data[["Lyapunov equation"]], "Lyapunov equation", first_three_limits, FALSE),
  focused_panel(panel_data[["Full edge message"]], "Full edge message", second_row_limits, TRUE),
  focused_panel(panel_data[["Full edge message gradient"]], "Full edge message gradient", second_row_limits, FALSE),
  focused_leg_panel(),
  design = "ABCD\nAEFG",
  widths = c(0.06, 1, 1, 1)
)

exp_adjoint_cached_fig
```

```{r exp-adjoint-cached-save}
save_benchmark_pdf(
  exp_adjoint_cached_fig,
  "ou_exp_adjoint_cached_benchmarks",
  width = 10,
  height = 5.8
)
```

```{r dt-stress-figure-removed, eval=FALSE}
dt_stress <- read_benchmark_csv(branch_results_path, "smbp_exp_dt_stress_results.csv")

if (!"seed" %in% names(dt_stress)) {
  dt_stress <- dt_stress %>%
    mutate(seed = 1L)
}

dt_stress_long <- dt_stress %>%
  transmute(
    seed,
    dt,
    `Dense exp` = dense_relerr,
    SMBP = smbp_relerr
  ) %>%
  pivot_longer(
    cols = c(`Dense exp`, SMBP),
    names_to = "method",
    values_to = "relative_error"
  ) %>%
  filter(is.finite(relative_error), relative_error > 0)

dt_stress_summary <- dt_stress_long %>%
  group_by(dt, method) %>%
  summarize(
    median_error = median(relative_error),
    q25 = quantile(relative_error, 0.25),
    q75 = quantile(relative_error, 0.75),
    .groups = "drop"
  )

dt_stress_arrows <- bind_rows(lapply(c(1.0e-2, 1.0), function(dt_value) {
  arrow_rows <- dt_stress_summary %>%
    filter(dplyr::near(dt, dt_value))
  dense_row <- arrow_rows %>%
    filter(method == "Dense exp")
  smbp_row <- arrow_rows %>%
    filter(method == "SMBP")
  if (nrow(dense_row) == 0L || nrow(smbp_row) == 0L) {
    return(tibble())
  }

  dense_error <- dense_row$median_error[[1L]]
  smbp_error <- smbp_row$median_error[[1L]]
  label_side <- if_else(dt_value < 1, "right", "left")
  arrow_gap <- 1.28
  tibble(
    dt = dt_value,
    y_start = dense_error / arrow_gap,
    y_end = smbp_error * arrow_gap,
    x_label = if_else(label_side == "right", dt_value * 1.45, dt_value / 1.45),
    y_label = sqrt(dense_error * smbp_error),
    hjust = if_else(label_side == "right", 0, 1),
    label = paste0(signif(dense_error / smbp_error, 2), "x")
  )
}))

dt_stress_colors <- c(
  "Dense exp" = kernel_colors[["Dense"]],
  "SMBP" = kernel_colors[["SSBP (R invertible, LDU)"]]
)

dt_stress_fig <- ggplot(
  dt_stress_summary,
  aes(dt, median_error, colour = method, linetype = method)
) +
  geom_ribbon(
    aes(ymin = q25, ymax = q75, fill = method),
    alpha = 0.12,
    linewidth = 0,
    show.legend = FALSE
  ) +
  geom_line(linewidth = 0.65) +
  geom_point(size = 1.8) +
  geom_segment(
    data = dt_stress_arrows,
    aes(x = dt, xend = dt, y = y_start, yend = y_end),
    inherit.aes = FALSE,
    colour = "black",
    linewidth = 0.3,
    arrow = arrow(type = "open", length = grid::unit(0.05, "inches"))
  ) +
  geom_text(
    data = dt_stress_arrows,
    aes(x = x_label, y = y_label, label = label, hjust = hjust),
    inherit.aes = FALSE,
    size = 3.2,
    family = fnt,
    colour = "black"
  ) +
  scale_x_log10(
    breaks = 10^(-4:3),
    labels = function(x) scales::math_format(10^.x)(log10(x))
  ) +
  scale_y_log10(
    breaks = 10^(-16:-10),
    labels = function(x) scales::math_format(10^.x)(log10(x))
  ) +
  scale_colour_manual(values = dt_stress_colors, name = NULL) +
  scale_fill_manual(values = dt_stress_colors, name = NULL) +
  scale_linetype_manual(values = c("Dense exp" = "32", "SMBP" = "solid"), name = NULL) +
  labs(
    x = "Transition scale (t)",
    y = "Relative error",
    title = "Fixed-K transition-scale stress test",
    subtitle = paste0(
      "K = ", unique(dt_stress$K),
      "; median over ", n_distinct(dt_stress$seed),
      " generated drifts"
    )
  ) +
  paper_theme(base_size = 9) +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.subtitle = element_text(size = 8)
  )

dt_stress_fig
```

```{r dt-stress-save, eval=FALSE}
save_benchmark_pdf(dt_stress_fig, "ou_dt_stress_benchmarks", width = 4.5, height = 3.2)
```

```{r stability-figure, fig.width=9.2, fig.height=6.4}
boundary_stability <- read_benchmark_csv(branch_results_path, "smbp_boundary_stability_results.csv")
slow_lyapunov <- read_benchmark_csv(branch_results_path, "smbp_slow_lyapunov_results.csv")
branch_covariance <- read_benchmark_csv(branch_results_path, "smbp_branch_covariance_stability_results.csv")
basis_conditioning <- read_benchmark_csv(branch_results_path, "smbp_basis_conditioning_results.csv")
lyapunov_param_stability <- read_benchmark_csv(branch_results_path, "lyapunov_param_stability_results.csv")

stability_exp_col <- kernel_colors[["SSBP (R invertible, LDU)"]]
stability_lyapunov_col <- kernel_colors[["Dense"]]
stability_branch_col <- kernel_colors[["Known Schur"]]

stability_cols <- c(
  "Exponential" = stability_exp_col,
  "Exp. adjoint" = stability_exp_col,
  "Lyapunov" = stability_lyapunov_col,
  "Lyap. adjoint" = stability_lyapunov_col,
  "Covariance error" = stability_branch_col,
  "Covariance adjoint" = stability_branch_col,
  "Relative covariance error" = stability_branch_col,
  "Absolute covariance error" = stability_branch_col,
  "Absolute branch-gradient error" = stability_branch_col,
  "Lyapunov residual" = stability_lyapunov_col,
  "Min. eigenvalue:\nbranch covariance" = "black",
  "Min. eigenvalue:\nstationary covariance" = "black",
  "Branch covariance" = stability_branch_col,
  "Branch gradient" = stability_branch_col
)

stability_linetypes <- c(
  "Exponential" = "solid",
  "Exp. adjoint" = "42",
  "Lyapunov" = "solid",
  "Lyap. adjoint" = "42",
  "Covariance error" = "solid",
  "Covariance adjoint" = "42",
  "Relative covariance error" = "solid",
  "Absolute covariance error" = "solid",
  "Absolute branch-gradient error" = "42",
  "Lyapunov residual" = "solid",
  "Min. eigenvalue:\nbranch covariance" = "solid",
  "Min. eigenvalue:\nstationary covariance" = "solid",
  "Branch covariance" = "solid",
  "Branch gradient" = "42"
)

log10_sparse_labels <- function(x) {
  pow <- log10(x)
  ifelse(
    abs(pow - round(pow)) < 1e-8 & round(pow) %% 2 == 0,
    scales::math_format(10^.x)(pow),
    ""
  )
}

log10_axis_labels <- function(x) {
  scales::math_format(10^.x)(log10(x))
}

precision_reference_layers <- function(include_sqrt = FALSE) {
  layers <- list(
    geom_hline(
      yintercept = .Machine$double.eps,
      colour = "black",
      linetype = "solid",
      linewidth = 0.2
    )
  )
  if (include_sqrt) {
    layers <- c(
      layers,
      list(
        geom_hline(
          yintercept = sqrt(.Machine$double.eps),
          colour = "grey78",
          linetype = "13",
          linewidth = 0.32
        )
      )
    )
  }
  layers
}

stability_panel_theme <- paper_theme(base_size = 8) +
  theme(
    legend.position = "bottom",
    legend.key.width = grid::unit(0.24, "inches"),
    legend.text = element_text(size = 6.2),
    panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold"),
    plot.margin = margin(2, 4, 2, 4, "pt")
  )

boundary_summary <- boundary_stability %>%
  { if (!"seed" %in% names(.)) mutate(., seed = 1L) else . } %>%
  transmute(
    seed,
    side,
    boundary_distance,
    Exponential = exp_relerr,
    `Exp. adjoint` = exp_adjoint_relerr,
    Lyapunov = lyapunov_relerr,
    `Lyap. adjoint` = lyapunov_adjoint_relerr,
    `Branch covariance` = branch_covariance_relerr,
    `Branch gradient` = branch_gradient_relerr
  ) %>%
  pivot_longer(
    cols = c(Exponential, `Exp. adjoint`, Lyapunov, `Lyap. adjoint`,
             `Branch covariance`, `Branch gradient`),
    names_to = "metric",
    values_to = "value"
  ) %>%
  filter(is.finite(value), value > 0) %>%
  group_by(boundary_distance, metric) %>%
  summarize(
    median_error = median(value),
    q25 = quantile(value, 0.25),
    q75 = quantile(value, 0.75),
    .groups = "drop"
  )

boundary_panel <- ggplot(
  boundary_summary,
  aes(boundary_distance, median_error, colour = metric, linetype = metric)
) +
  # geom_ribbon(
  #   aes(ymin = q25, ymax = q75, fill = metric),
  #   alpha = 0.12,
  #   linewidth = 0,
  #   show.legend = FALSE
  # ) +
  geom_line(linewidth = 0.55) +
  scale_x_log10(labels = log10_axis_labels) +
  scale_y_log10(
    breaks = 10^(-17:-13),
    limits = c(1.0e-17, 1.0e-13),
    labels = log10_axis_labels
  ) +
  scale_colour_manual(values = stability_cols, name = NULL) +
  scale_fill_manual(values = stability_cols, name = NULL) +
  scale_linetype_manual(values = stability_linetypes, name = NULL) +
  labs(
    x = "Distance to boundary",
    y = "Relative error",
    title = "Repeated-root boundary"
  ) +
  precision_reference_layers(include_sqrt = FALSE) +
  stability_panel_theme

slow_panel_error_data <- slow_lyapunov %>%
  transmute(
    epsilon,
    `Covariance error` = sigma_relerr,
    `Covariance adjoint` = lyapunov_adjoint_relerr,
    `Lyapunov residual` = residual_relerr
  ) %>%
  pivot_longer(
    cols = -epsilon,
    names_to = "metric",
    values_to = "value"
  ) %>%
  filter(is.finite(value), value > 0)

slow_eigen_axis_scale <- sqrt(.Machine$double.eps) /
  max(slow_lyapunov$min_eigenvalue, na.rm = TRUE)

slow_panel_eigen_data <- slow_lyapunov %>%
  transmute(
    epsilon,
    metric = "Min. eigenvalue:\nstationary covariance",
    value = pmax(min_eigenvalue, .Machine$double.xmin) * slow_eigen_axis_scale
  ) %>%
  filter(is.finite(value), value > 0)

slow_panel <- ggplot(
  slow_panel_error_data,
  aes(epsilon, value, colour = metric, linetype = metric)
) +
  geom_line(linewidth = 0.55) +
  geom_line(
    data = slow_panel_eigen_data,
    linewidth = 0.55
  ) +
  scale_x_log10(labels = log10_axis_labels) +
  scale_y_log10(
    labels = log10_axis_labels,
    sec.axis = sec_axis(
      ~ . / slow_eigen_axis_scale,
      name = "Min. eigenvalue of stationary covariance",
      labels = log10_axis_labels
    )
  ) +
  scale_colour_manual(values = stability_cols, name = NULL) +
  scale_linetype_manual(values = stability_linetypes, name = NULL) +
  labs(
    x = "Mean-reversion scale",
    y = "Relative error",
    title = "Slow mean reversion"
  ) +
  precision_reference_layers(include_sqrt = FALSE) +
  stability_panel_theme

branch_panel_data <- branch_covariance %>%
  transmute(
    dt,
    `Relative covariance error` = q_relerr,
    `Absolute covariance error` = q_abs_fro_error,
    `Absolute branch-gradient error` = branch_gradient_abs_fro_error,
    `Min. eigenvalue:\nbranch covariance` = pmax(min_eigenvalue, .Machine$double.xmin)
  ) %>%
  pivot_longer(
    cols = -dt,
    names_to = "metric",
    values_to = "value"
  ) %>%
  filter(is.finite(value), value > 0)

branch_panel <- ggplot(
  branch_panel_data,
  aes(dt, value, colour = metric, linetype = metric)
) +
  geom_line(linewidth = 0.55) +
  scale_x_log10(labels = log10_axis_labels) +
  scale_y_log10(labels = log10_axis_labels) +
  scale_colour_manual(
    values = c(
      "Relative covariance error" = stability_cols[["Relative covariance error"]],
      "Absolute covariance error" = stability_cols[["Absolute covariance error"]],
      "Absolute branch-gradient error" = stability_cols[["Absolute branch-gradient error"]],
      "Min. eigenvalue:\nbranch covariance" =
        stability_cols[["Min. eigenvalue:\nbranch covariance"]]
    ),
    name = NULL
  ) +
  scale_linetype_manual(values = stability_linetypes, name = NULL) +
  labs(
    x = "Branch length",
    y = "Error or branch-covariance eigenvalue",
    title = "Innovation covariance"
  ) +
  precision_reference_layers(include_sqrt = FALSE) +
  stability_panel_theme

conditioning_panel_data <- basis_conditioning %>%
  transmute(
    observed_condition,
    Exponential = exp_relerr,
    `Exp. adjoint` = exp_adjoint_relerr,
    Lyapunov = lyapunov_relerr,
    `Lyap. adjoint` = lyapunov_adjoint_relerr,
    `Branch covariance` = branch_covariance_relerr,
    `Branch gradient` = branch_gradient_relerr
  ) %>%
  pivot_longer(
    cols = -observed_condition,
    names_to = "metric",
    values_to = "value"
  ) %>%
  filter(is.finite(value), value > 0)

conditioning_panel <- ggplot(
  conditioning_panel_data,
  aes(observed_condition, value, colour = metric, linetype = metric)
) +
  geom_line(linewidth = 0.55) +
  scale_x_log10(labels = log10_axis_labels) +
  scale_y_log10(labels = log10_axis_labels) +
  scale_colour_manual(
    values = c(
      "Exponential" = stability_cols[["Exponential"]],
      "Exp. adjoint" = stability_cols[["Exp. adjoint"]],
      "Lyapunov" = stability_cols[["Lyapunov"]],
      "Lyap. adjoint" = stability_cols[["Lyap. adjoint"]],
      "Branch covariance" = stability_cols[["Branch covariance"]],
      "Branch gradient" = stability_cols[["Branch covariance"]]
    ),
    name = NULL
  ) +
  scale_linetype_manual(values = stability_linetypes, name = NULL) +
  labs(
    x = "Condition number",
    y = "Relative error",
    title = "Change-of-basis conditioning"
  ) +
  precision_reference_layers(include_sqrt = TRUE) +
  stability_panel_theme

stability_fig <- (boundary_panel | conditioning_panel) / (branch_panel | slow_panel) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 9, family = "serif"))

stability_fig
```

```{r stability-save, eval=FALSE}
save_benchmark_pdf(stability_fig, "ou_numerical_stability_benchmarks", width = 9.2, height = 6.4)
```

```{r lyapunov-param-stability-figure, fig.width=9.2, fig.height=6.4}
lyapunov_param_metric_cols <- c(
  "Matrix exponential" = stability_cols[["Exponential"]],
  "Branch covariance" = stability_cols[["Branch covariance"]],
  "Absolute branch covariance" = "#6F7B82",
  "Lyapunov identity residual" = stability_cols[["Lyapunov residual"]]
)

lyapunov_param_metric_linetypes <- c(
  "Matrix exponential" = "solid",
  "Branch covariance" = "solid",
  "Absolute branch covariance" = "42",
  "Lyapunov identity residual" = "32"
)

lyapunov_param_plot_data <- lyapunov_param_stability %>%
  transmute(
    stress,
    x_value,
    `Matrix exponential` = exp_relerr,
    `Branch covariance` = branch_covariance_relerr,
    `Absolute branch covariance` = branch_covariance_abs_fro_error,
    `Lyapunov identity residual` = lyapunov_residual_relerr
  ) %>%
  pivot_longer(
    cols = -c(stress, x_value),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(
    metric = factor(metric, levels = names(lyapunov_param_metric_cols))
  ) %>%
  filter(is.finite(value), value > 0)

lyapunov_param_panel <- function(stress_value, title, x_label) {
  panel_data <- lyapunov_param_plot_data %>%
    filter(stress == stress_value)

  ggplot(
    panel_data,
    aes(x_value, value, colour = metric, linetype = metric)
  ) +
    geom_line(linewidth = 0.55) +
    scale_x_log10(labels = log10_axis_labels) +
    scale_y_log10(
      breaks = 10^(-18:0),
      labels = log10_sparse_labels,
      minor_breaks = NULL
    ) +
    scale_colour_manual(values = lyapunov_param_metric_cols, name = NULL) +
    scale_linetype_manual(values = lyapunov_param_metric_linetypes, name = NULL) +
    labs(
      x = x_label,
      y = "Error",
      title = title
    ) +
    precision_reference_layers(include_sqrt = FALSE) +
    stability_panel_theme
}

lyapunov_param_stability_fig <- (
  lyapunov_param_panel(
    "stationary-covariance-conditioning",
    "Stationary covariance conditioning",
    expression(kappa(Sigma))
  ) |
    lyapunov_param_panel(
      "diffusion-conditioning",
      "Diffusion conditioning",
      expression(kappa(Q))
    )
) / (
  lyapunov_param_panel(
    "skew-strength",
    "Skew-symmetric strength",
    expression(gamma)
  ) |
    lyapunov_param_panel(
      "slow-mean-reversion",
      "Slow mean reversion",
      expression(epsilon)
    )
) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(face = "bold", size = 9, family = "serif")
  )

lyapunov_param_stability_fig
```

```{r lyapunov-param-stability-save, eval=FALSE}
save_benchmark_pdf(
  lyapunov_param_stability_fig,
  "ou_lyapunov_param_stability_benchmarks",
  width = 9.2,
  height = 6.4
)
```

```{r ssbp-vs-lyapunov-param-stability-figure, fig.width=9.2, fig.height=7.0}
combo_relabel <- function(metric) {
  dplyr::recode(
    as.character(metric),
    "Exponential" = "Matrix exponential",
    "Lyapunov" = "Lyapunov solve/identity",
    "Lyapunov identity residual" = "Lyapunov solve/identity",
    .default = as.character(metric)
  )
}

combo_boundary_data <- boundary_summary %>%
  mutate(metric = combo_relabel(metric))

combo_conditioning_data <- conditioning_panel_data %>%
  mutate(metric = combo_relabel(metric))

combo_lyapunov_param_data <- lyapunov_param_plot_data %>%
  filter(metric != "Absolute branch covariance") %>%
  mutate(metric = combo_relabel(metric))

# Ordered so that a 2-column, row-major legend (see ssbp_boundary_combo_panel
# below) pairs each forward quantity with its adjoint on the same row.
combo_cols <- c(
  "Matrix exponential" = lyapunov_param_metric_cols[["Matrix exponential"]],
  "Exp. adjoint" = stability_cols[["Exp. adjoint"]],
  "Lyapunov solve/identity" = stability_lyapunov_col,
  "Lyap. adjoint" = stability_cols[["Lyap. adjoint"]],
  "Branch covariance" = stability_cols[["Branch covariance"]],
  "Branch gradient" = stability_cols[["Branch covariance"]]
)

combo_linetypes <- c(
  "Matrix exponential" = "solid",
  "Exp. adjoint" = "42",
  "Lyapunov solve/identity" = "solid",
  "Lyap. adjoint" = "42",
  "Branch covariance" = "solid",
  "Branch gradient" = "42"
)

# One shared y-axis range across both rows (not just within a row), per
# request: the SSBP and Lyapunov-param panels must show the same range.
combo_y_breaks <- 10^(-18:0)
combo_y_limits <- c(1.0e-17, 1.0e-6)

combo_hide_y_axis <- theme(
  axis.title.y = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank()
)

# Scale text elements so they land at specific printed sizes once this
# 9.2in-wide figure is placed at 0.98\textwidth in a 6.5in-textwidth, 12pt
# document (printed pt = raw ggplot pt * combo_scale_down). Built from
# paper_theme()/stability_panel_theme() rather than mutating them, since
# those are shared by other figures in this file.
combo_scale_down <- 0.98 * 6.5 / 9.2
combo_font_scale <- (12 / combo_scale_down) / 9
# Panel/row titles are bold but otherwise match the (non-bold) axis-title
# size exactly, since both are scaled in raw ggplot units, their printed
# sizes after the same width-based embed scaling are equal too.
combo_title_pt <- 8 * combo_font_scale
combo_panel_theme <- paper_theme(base_size = 8 * combo_font_scale) +
  theme(
    legend.position = "bottom",
    legend.key.width = grid::unit(0.24, "inches"),
    legend.text = element_text(size = 6.2 * combo_font_scale),
    panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, size = combo_title_pt, face = "bold"),
    plot.margin = margin(2, 4, 2, 4, "pt")
  )

ssbp_boundary_combo_panel <- ggplot(
  combo_boundary_data,
  aes(boundary_distance, median_error, colour = metric, linetype = metric)
) +
  geom_line(linewidth = 0.55) +
  scale_x_log10(labels = log10_axis_labels) +
  scale_y_log10(
    breaks = combo_y_breaks,
    limits = combo_y_limits,
    labels = log10_sparse_labels,
    minor_breaks = NULL
  ) +
  scale_colour_manual(values = combo_cols, limits = names(combo_cols), name = NULL) +
  scale_linetype_manual(values = combo_linetypes, limits = names(combo_linetypes), name = NULL) +
  guides(
    colour = guide_legend(ncol = 2, byrow = TRUE),
    linetype = guide_legend(ncol = 2, byrow = TRUE)
  ) +
  labs(x = "Distance to boundary", y = NULL, title = "Repeated-root boundary") +
  precision_reference_layers(include_sqrt = FALSE) +
  combo_panel_theme +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.02, 0.98),
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.background = element_rect(fill = "white", colour = "black", linewidth = 0.3),
    legend.margin = margin(3, 5, 3, 4, "pt"),
    legend.spacing.x = grid::unit(2, "pt"),
    legend.spacing.y = grid::unit(0, "pt"),
    legend.key.size = grid::unit(11, "pt"),
    legend.key.spacing.x = grid::unit(3, "pt"),
    legend.key.spacing.y = grid::unit(1, "pt")
  )

ssbp_conditioning_combo_panel <- ggplot(
  combo_conditioning_data,
  aes(observed_condition, value, colour = metric, linetype = metric)
) +
  geom_line(linewidth = 0.55) +
  scale_x_log10(labels = log10_axis_labels) +
  scale_y_log10(
    breaks = combo_y_breaks,
    limits = combo_y_limits,
    labels = log10_sparse_labels,
    minor_breaks = NULL
  ) +
  scale_colour_manual(values = combo_cols, limits = names(combo_cols), name = NULL) +
  scale_linetype_manual(values = combo_linetypes, limits = names(combo_linetypes), name = NULL) +
  labs(
    x = expression(paste("Condition number ", kappa, "(", bold(R), ")")),
    y = NULL,
    title = "Change-of-basis conditioning"
  ) +
  precision_reference_layers(include_sqrt = TRUE) +
  annotate(
    "segment",
    x = 1.0e3, y = 4.0e-11, xend = 1.0e5, yend = 1.0e-12,
    arrow = arrow(length = grid::unit(0.09, "inches"), type = "closed"),
    linewidth = 0.35, colour = "black"
  ) +
  annotate(
    "text",
    x = 2.2, y = 1.0e-10,
    label = expression(paste("Solved when ", bold(R)^{-1}, " = ", bold(R)^{T})),
    hjust = 0, vjust = 0,
    size = (0.8 * 8 * combo_font_scale) / 2.845,
    family = fnt, colour = "black"
  ) +
  combo_panel_theme +
  theme(legend.position = "none") +
  combo_hide_y_axis

lyapunov_param_combo_panel <- function(stress_value, title, x_label, hide_y_axis = FALSE) {
  panel_data <- combo_lyapunov_param_data %>%
    filter(stress == stress_value)

  p <- ggplot(
    panel_data,
    aes(x_value, value, colour = metric, linetype = metric)
  ) +
    geom_line(linewidth = 0.55) +
    scale_x_log10(labels = log10_axis_labels) +
    scale_y_log10(
      breaks = combo_y_breaks,
      limits = combo_y_limits,
      labels = log10_sparse_labels,
      minor_breaks = NULL
    ) +
    scale_colour_manual(values = combo_cols, limits = names(combo_cols), name = NULL) +
    scale_linetype_manual(values = combo_linetypes, limits = names(combo_linetypes), name = NULL) +
    labs(x = x_label, y = NULL, title = title) +
    precision_reference_layers(include_sqrt = FALSE) +
    combo_panel_theme +
    theme(legend.position = "none")

  if (hide_y_axis) p <- p + combo_hide_y_axis
  p
}

combo_rotated_label_panel <- function(lines, size_mm, bold = TRUE) {
  # Each line of text is its own rotated annotate() call, placed side by
  # side; a single annotate() call with an embedded "\n" renders its second
  # line mirrored under angle = 90, so multi-line labels are built this way
  # instead.
  n <- length(lines)
  gap <- 1.4
  x_pos <- seq(-(n - 1) / 2, (n - 1) / 2, length.out = n) * gap
  x_limit <- max(1, max(abs(x_pos)) + 0.7)

  p <- ggplot() +
    scale_x_continuous(limits = c(-x_limit, x_limit)) +
    scale_y_continuous(limits = c(-1, 1)) +
    theme_void()
  for (i in seq_len(n)) {
    p <- p + annotate(
      "text", x = x_pos[i], y = 0, label = lines[i],
      angle = 90,
      fontface = if (bold) "bold" else "plain",
      size = size_mm, family = fnt, colour = "black"
    )
  }
  p
}

combo_row_label_panel <- function(lines) {
  combo_rotated_label_panel(lines, size_mm = combo_title_pt / 2.845, bold = TRUE)
}

# Matches the (non-bold) size of the other axis titles, in mm (ggplot's
# annotate("text", size=) unit) rather than the pt-based theme element size.
combo_y_axis_label_panel <- function(label) {
  combo_rotated_label_panel(label, size_mm = (8 * combo_font_scale) / 2.845, bold = FALSE)
}

ssbp_vs_lyapunov_param_fig <- wrap_plots(
  combo_row_label_panel("H-SSBP"),
  ssbp_boundary_combo_panel,
  ssbp_conditioning_combo_panel,
  combo_row_label_panel("Lyap.-based para."),
  lyapunov_param_combo_panel(
    "skew-strength",
    "Skew-symmetric strength",
    expression(paste("Rotational strength ", bold("||M||"))),
    hide_y_axis = FALSE
  ),
  lyapunov_param_combo_panel(
    "stationary-covariance-conditioning",
    "Stationary covariance conditioning",
    expression(paste("Conditioning number ", kappa, "(", bold(V), ")")),
    hide_y_axis = TRUE
  ),
  combo_y_axis_label_panel("Relative error"),
  design = "YABC\nYDEF",
  widths = c(0.05, 0.12, 1, 1),
  heights = c(1, 1)
)

ssbp_vs_lyapunov_param_fig
```

```{r ssbp-vs-lyapunov-param-stability-save}
save_benchmark_pdf(
  ssbp_vs_lyapunov_param_fig,
  "ou_ssbp_vs_lyapunov_param_stability_summary",
  width = 9.2,
  height = 7.0
)
```

```{r matched-param-stability-figure, eval=FALSE}
matched_param_method_linetypes <- c(
  "H-SSBP" = "solid",
  "Lyapunov parametrization" = "dashed"
)

matched_param_metric_cols <- c(
  "Drift reconstruction" = kernel_colors[["SSBP (R orthogonal)"]],
  "Matrix exponential" = stability_cols[["Exponential"]],
  "Branch covariance" = stability_cols[["Branch covariance"]],
  "Lyapunov identity residual" = stability_cols[["Lyapunov residual"]]
)

matched_param_plot_data <- matched_param_stability %>%
  transmute(
    stress,
    group,
    x_value,
    method,
    `Drift reconstruction` = drift_relerr,
    `Matrix exponential` = exp_relerr,
    `Branch covariance` = branch_covariance_relerr,
    `Lyapunov identity residual` = lyapunov_residual_relerr
  ) %>%
  pivot_longer(
    cols = -c(stress, group, x_value, method),
    names_to = "metric",
    values_to = "value"
  ) %>%
  filter(is.finite(value), value > 0) %>%
  group_by(stress, x_value, method, metric) %>%
  summarize(value = median(value), .groups = "drop") %>%
  mutate(
    method = recode(method, "SSBP" = "H-SSBP"),
    method = factor(method, levels = names(matched_param_method_linetypes)),
    metric = factor(metric, levels = names(matched_param_metric_cols))
  )

matched_param_panel <- function(stress_value, title, x_label, show_y = TRUE) {
  panel_data <- matched_param_plot_data %>%
    filter(stress == stress_value)

  plot <- ggplot(
    panel_data,
    aes(x_value, value, colour = metric, linetype = method)
  ) +
    geom_line(linewidth = 0.52) +
    scale_x_log10(labels = log10_axis_labels) +
    scale_y_log10(
      limits = c(1.0e-18, 1.0),
      breaks = 10^seq(-18, 0, by = 2),
      labels = log10_sparse_labels,
      minor_breaks = NULL
    ) +
    scale_colour_manual(values = matched_param_metric_cols, name = NULL) +
    scale_linetype_manual(values = matched_param_method_linetypes, name = NULL) +
    labs(
      x = x_label,
      y = if (show_y) "Relative error" else NULL,
      title = title
    ) +
    precision_reference_layers(include_sqrt = FALSE) +
    stability_panel_theme

  if (!show_y) {
    plot <- plot +
      theme(
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
      )
  }
  plot
}

matched_param_stability_fig <- (
  matched_param_panel(
    "repeated-root-boundary",
    "Repeated-root boundary",
    "Distance to boundary",
    show_y = TRUE
  ) |
    matched_param_panel(
      "basis-conditioning",
      "Matched basis conditioning",
      expression(kappa(R)),
      show_y = FALSE
    )
) / (
  matched_param_panel(
    "branch-length",
    "Matched branch length",
    "Branch length",
    show_y = TRUE
  ) |
    matched_param_panel(
      "slow-mean-reversion",
      "Matched slow mean reversion",
      expression(epsilon),
      show_y = FALSE
    )
) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(
    legend.position = "bottom",
    plot.tag = element_text(face = "bold", size = 9, family = "serif")
  )

matched_param_stability_fig
```

```{r matched-param-stability-save, eval=FALSE}
save_benchmark_pdf(
  matched_param_stability_fig,
  "ou_matched_param_stability_benchmarks",
  width = 9.2,
  height = 6.4
)
```

```{r gradient-stress-table, eval=FALSE}
tex_sci <- function(x) {
  if (!is.finite(x)) {
    return("--")
  }
  if (x == 0) {
    return("$0$")
  }
  exponent <- floor(log10(abs(x)))
  mantissa <- x / 10^exponent
  paste0("$", sprintf("%.2g", mantissa), "\\times 10^{", exponent, "}$")
}

gradient_table <- gradient_stress %>%
  mutate(
    regime = recode(
      regime,
      "well-conditioned" = "Well-conditioned",
      "near-boundary" = "Near boundary",
      "slow-mean-reversion" = "Slow mean reversion",
      "ill-conditioned-basis" = "Ill-conditioned basis",
      "small-branch-length" = "Small branch length"
    ),
    chol = if_else(branch_cholesky_success, "yes", "no")
  )

gradient_table_lines <- c(
  "\\begin{tabular}{lcccc}",
  "\\toprule",
  "Regime & Exp. adjoint & Lyap. adjoint & Branch gradient & Cholesky \\\\",
  "\\midrule",
  vapply(seq_len(nrow(gradient_table)), function(i) {
    paste(
      gradient_table$regime[[i]],
      tex_sci(gradient_table$exp_adjoint_relerr[[i]]),
      tex_sci(gradient_table$lyapunov_adjoint_relerr[[i]]),
      tex_sci(gradient_table$branch_gradient_relerr[[i]]),
      gradient_table$chol[[i]],
      sep = " & "
    ) |>
      paste0(" \\\\")
  }, character(1)),
  "\\bottomrule",
  "\\end{tabular}"
)

save_benchmark_text(gradient_table_lines, "smbp_gradient_stress_table.tex")
```

```{r gradient-stress-figure, eval=FALSE}
gradient_regime_levels <- c(
  "well-conditioned",
  "near-boundary",
  "slow-mean-reversion",
  "ill-conditioned-basis",
  "small-branch-length"
)

gradient_regime_labels <- c(
  "Well-\nconditioned",
  "Near\nboundary",
  "Slow mean\nreversion",
  "Ill-conditioned\nbasis",
  "Small branch\nlength"
)

gradient_metric_cols <- c(
  "Exp. adjoint" = stability_cols[["Exp. adjoint"]],
  "Lyap. adjoint" = stability_cols[["Lyap. adjoint"]],
  "Branch gradient" = stability_cols[["Branch gradient"]]
)

gradient_plot_data <- gradient_stress %>%
  mutate(
    regime = factor(
      regime,
      levels = gradient_regime_levels,
      labels = gradient_regime_labels
    )
  ) %>%
  transmute(
    regime,
    `Exp. adjoint` = exp_adjoint_relerr,
    `Lyap. adjoint` = lyapunov_adjoint_relerr,
    `Branch gradient` = branch_gradient_relerr
  ) %>%
  pivot_longer(
    cols = -regime,
    names_to = "metric",
    values_to = "relative_error"
  ) %>%
  mutate(
    metric = factor(
      metric,
      levels = c("Exp. adjoint", "Lyap. adjoint", "Branch gradient")
    )
  ) %>%
  filter(is.finite(relative_error), relative_error > 0)

gradient_stress_fig <- ggplot(
  gradient_plot_data,
  aes(regime, relative_error, colour = metric, shape = metric)
) +
  geom_point(
    position = position_dodge(width = 0.52),
    size = 2.15,
    stroke = 0.35
  ) +
  scale_y_log10(
    breaks = 10^seq(-17, -2, by = 3),
    limits = c(1.0e-17, 1.0e-2),
    labels = log10_axis_labels
  ) +
  scale_colour_manual(values = gradient_metric_cols, name = NULL) +
  scale_shape_manual(
    values = c("Exp. adjoint" = 16, "Lyap. adjoint" = 17, "Branch gradient" = 15),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = "Relative error",
    title = "Gradient stress"
  ) +
  paper_theme(base_size = 8.5) +
  theme(
    legend.position = "bottom",
    legend.key.width = grid::unit(0.38, "inches"),
    legend.text = element_text(size = 7.5),
    panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 7.5),
    plot.title = element_text(hjust = 0.5, size = 9.5, face = "bold")
  )

gradient_stress_fig
```

```{r gradient-stress-save, eval=FALSE}
save_benchmark_pdf(gradient_stress_fig, "ou_gradient_stress_benchmarks", width = 5.8, height = 3.2)
```

# Time-Series and Tree Panels

The panel code below follows the benchmark plotting framework used in the
CTMC benchmark figure. The time-series panel reads the available CSV files and
fills missing cells by cubic extrapolation from the nearest observed anchor.
The tree panel reads package benchmark CSV files, estimates missing package cells
the same way, and estimates the SMB tree curve from the branch-with-Lyapunov
kernel benchmark multiplied by the number of tree edges.

```{r panel-data, eval=FALSE}
panel_k <- c(4L, 8L, 16L, 32L, 64L)
panel_n <- c(10L, 100L, 1000L)

read_existing_csvs <- function(paths) {
  paths <- paths[fs::file_exists(paths)]
  if (length(paths) == 0L) {
    return(tibble())
  }
  bind_rows(lapply(paths, readr::read_csv, show_col_types = FALSE))
}

time_series_grid_files <- fs::dir_ls(
  time_series_results_path,
  regexp = "timeseries_.*\\.csv$",
  fail = FALSE
)

time_series_observed <- read_existing_csvs(time_series_grid_files) %>%
  filter(
    status == "ok",
    operation %in% c("log_prob", "grad_log_prob"),
    K %in% panel_k,
    N %in% panel_n
  ) %>%
  mutate(
    operation = recode(
      operation,
      log_prob = "Likelihood",
      grad_log_prob = "Gradient"
    ),
    method = case_when(
      package == "rstan" ~ "Stan",
      str_detect(str_to_lower(package), "beast|smb") ~ "SMB",
      TRUE ~ as.character(package)
    ),
    median_s = median_ms / 1e3,
    sd_s = pmax(max_ms - min_ms, 0) / (4 * 1e3),
    is_extrap = FALSE
  ) %>%
  select(method, operation, N, K, median_s, sd_s, is_extrap) %>%
  distinct(method, operation, N, K, .keep_all = TRUE)

complete_cubic_time_series_grid <- function(
  data,
  all_N = panel_n,
  all_K = panel_k,
  k_power = 3,
  n_power = 3
) {
  if (nrow(data) == 0L) {
    return(data)
  }

  out <- list()
  keys <- distinct(data, method, operation)
  row_id <- 1L

  for (i in seq_len(nrow(keys))) {
    method_i <- as.character(keys$method[[i]])
    operation_i <- as.character(keys$operation[[i]])
    source_data <- data %>%
      filter(method == method_i, operation == operation_i) %>%
      mutate(N = as.integer(N), K = as.integer(K))
    observed_N <- sort(unique(source_data$N))

    completed_observed_N <- list()
    for (n_value in observed_N) {
      n_data <- source_data %>% filter(N == n_value)
      completed_k <- lapply(all_K, function(k_value) {
        exact <- n_data %>% filter(K == k_value)
        if (nrow(exact) > 0L) {
          return(exact[1, ])
        }

        lower <- n_data %>% filter(K < k_value) %>% arrange(K)
        anchor <- if (nrow(lower) > 0L) {
          tail(lower, 1)
        } else {
          n_data %>% arrange(K) %>% slice(1)
        }
        scale <- (k_value / anchor$K[[1]])^k_power
        anchor %>%
          mutate(
            K = as.integer(k_value),
            median_s = median_s * scale,
            sd_s = sd_s * scale,
            is_extrap = TRUE
          )
      }) %>%
        bind_rows() %>%
        arrange(K)

      completed_observed_N[[as.character(n_value)]] <- completed_k
    }

    for (n_value in all_N) {
      n_key <- as.character(n_value)
      if (n_key %in% names(completed_observed_N)) {
        out[[row_id]] <- completed_observed_N[[n_key]]
        row_id <- row_id + 1L
        next
      }

      lower_N <- observed_N[observed_N < n_value]
      source_N <- if (length(lower_N) > 0L) {
        max(lower_N)
      } else {
        observed_N[[1L]]
      }
      scale <- (n_value / source_N)^n_power
      out[[row_id]] <- completed_observed_N[[as.character(source_N)]] %>%
        mutate(
          N = as.integer(n_value),
          median_s = median_s * scale,
          sd_s = sd_s * scale,
          is_extrap = TRUE
        )
      row_id <- row_id + 1L
    }
  }

  bind_rows(out) %>%
    arrange(operation, N, method, K)
}

complete_cubic_panel_grid <- function(
  data,
  all_N = panel_n,
  all_K = panel_k,
  k_power = 3,
  n_power = 3
) {
  if (nrow(data) == 0L) {
    return(data)
  }

  out <- list()
  row_id <- 1L
  for (method_i in unique(as.character(data$method))) {
    source_data <- data %>%
      filter(method == method_i) %>%
      mutate(N = as.integer(N), K = as.integer(K))
    observed_N <- sort(unique(source_data$N))

    completed_observed_N <- list()
    for (n_value in observed_N) {
      n_data <- source_data %>% filter(N == n_value)
      completed_k <- lapply(all_K, function(k_value) {
        exact <- n_data %>% filter(K == k_value)
        if (nrow(exact) > 0L) {
          return(exact[1, ])
        }

        lower <- n_data %>% filter(K < k_value) %>% arrange(K)
        anchor <- if (nrow(lower) > 0L) {
          tail(lower, 1)
        } else {
          n_data %>% arrange(K) %>% slice(1)
        }
        scale <- (k_value / anchor$K[[1]])^k_power
        anchor %>%
          mutate(
            K = as.integer(k_value),
            median_s = median_s * scale,
            sd_s = sd_s * scale,
            is_extrap = TRUE
          )
      }) %>%
        bind_rows() %>%
        arrange(K)

      completed_observed_N[[as.character(n_value)]] <- completed_k
    }

    for (n_value in all_N) {
      n_key <- as.character(n_value)
      if (n_key %in% names(completed_observed_N)) {
        out[[row_id]] <- completed_observed_N[[n_key]]
        row_id <- row_id + 1L
        next
      }

      lower_N <- observed_N[observed_N < n_value]
      source_N <- if (length(lower_N) > 0L) {
        max(lower_N)
      } else {
        observed_N[[1L]]
      }
      scale <- (n_value / source_N)^n_power
      out[[row_id]] <- completed_observed_N[[as.character(source_N)]] %>%
        mutate(
          N = as.integer(n_value),
          median_s = median_s * scale,
          sd_s = sd_s * scale,
          is_extrap = TRUE
        )
      row_id <- row_id + 1L
    }
  }

  bind_rows(out) %>%
    arrange(N, method, K)
}

tree_result_files <- fs::dir_ls(
  tree_results_path,
  regexp = "results_.*\\.csv$",
  fail = FALSE
)

tree_raw <- bind_rows(lapply(tree_result_files, function(path) {
  readr::read_csv(path, show_col_types = FALSE) %>%
    mutate(source_file = fs::path_file(path))
}))

tree_package_observed <- tree_raw %>%
  filter(
    status == "ok",
    K %in% panel_k,
    N %in% panel_n
  ) %>%
  mutate(
    method = case_when(
      package %in% c("PCMBase", "PCMBaseCpp") & mode == "dense_PCMLik" ~ "PCMBase",
      package == "PCMFit" & mode == "PCMFit_fit" ~ "PCMFit",
      package == "mvMORPH" & str_detect(mode, "^mvOU_fixed") ~ "mvMORPH",
      package == "mvSLOUCH" & mode == "ouchModel_fit_Any_Any" ~ "mvSLOUCH",
      TRUE ~ NA_character_
    ),
    median_s = case_when(
      is.finite(median_sec) & median_sec > 0 ~ median_sec,
      is.finite(mean_sec) & mean_sec > 0 ~ mean_sec,
      TRUE ~ NA_real_
    ),
    sd_s = pmax(max_sec - min_sec, 0) / 4,
    is_extrap = FALSE,
    package_priority = case_when(
      package == "PCMBaseCpp" ~ 1L,
      TRUE ~ 2L
    ),
    source_priority = case_when(
      str_detect(source_file, "_grid_") & !str_detect(source_file, "_quick_") ~ 1L,
      str_detect(source_file, "_quick_grid_") ~ 2L,
      TRUE ~ 3L
    )
  ) %>%
  filter(!is.na(method), is.finite(median_s), median_s > 0) %>%
  arrange(method, N, K, package_priority, source_priority, median_s) %>%
  group_by(method, N, K) %>%
  slice(1L) %>%
  ungroup() %>%
  select(method, N, K, median_s, sd_s, is_extrap)

branch_lyapunov_file <- fs::path(
  branch_results_path, "branch_likelihood_with_lyapunov_results.csv"
)
tree_smbp_estimated <- if (fs::file_exists(branch_lyapunov_file)) {
  readr::read_csv(branch_lyapunov_file, show_col_types = FALSE) %>%
    filter(K %in% panel_k, is.finite(smbp_branch_lyap_rotation_ms)) %>%
    select(K, smbp_branch_lyap_rotation_ms) %>%
    distinct(K, .keep_all = TRUE) %>%
    tidyr::crossing(N = panel_n) %>%
    transmute(
      method = "SMB",
      N = as.integer(N),
      K = as.integer(K),
      median_s = (2 * N - 2) * smbp_branch_lyap_rotation_ms / 1e3,
      sd_s = 0.10 * median_s,
      is_extrap = TRUE
    )
} else {
  tibble()
}

time_series_timings <- complete_cubic_time_series_grid(time_series_observed) %>%
  mutate(
    operation = factor(operation, levels = c("Likelihood", "Gradient")),
    method = factor(method, levels = c("SMB", "Stan", "Dense Kalman", "Known Schur"))
  )

tree_timings <- bind_rows(
  tree_smbp_estimated,
  complete_cubic_panel_grid(tree_package_observed)
) %>%
  mutate(method = factor(method, levels = c("SMB", "PCMBase", "PCMFit", "mvMORPH", "mvSLOUCH")))
```

```{r panel-framework, eval=FALSE}
fnt <- "CMU Serif"
fsz <- 10
fszT <- 12

method_cols <- c(
  "SMB" = "#2F67B1",
  "Stan" = "#D4960A",
  "Dense Kalman" = "#9B4356",
  "Known Schur" = "#D67C2F",
  "PCMBase" = "#D4960A",
  "PCMFit" = "#6C5B9A",
  "mvMORPH" = "#9B4356",
  "mvSLOUCH" = "#D67C2F"
)
method_shps <- c(
  "SMB" = 16L,
  "Stan" = 16L,
  "Dense Kalman" = 16L,
  "Known Schur" = 16L,
  "PCMBase" = 16L,
  "PCMFit" = 16L,
  "mvMORPH" = 16L,
  "mvSLOUCH" = 16L
)
method_alpha <- c(
  "SMB" = 1,
  "Stan" = 1,
  "Dense Kalman" = 1,
  "Known Schur" = 1,
  "PCMBase" = 1,
  "PCMFit" = 1,
  "mvMORPH" = 1,
  "mvSLOUCH" = 1
)
method_lty <- c(
  "SMB" = "solid",
  "Stan" = "dashed",
  "Dense Kalman" = "dashed",
  "Known Schur" = "dashed",
  "PCMBase" = "dashed",
  "PCMFit" = "dashed",
  "mvMORPH" = "dashed",
  "mvSLOUCH" = "dashed"
)
method_ltyLegend <- c(
  "SMB" = "solid",
  "Stan" = "32",
  "Dense Kalman" = "32",
  "Known Schur" = "32",
  "PCMBase" = "32",
  "PCMFit" = "32",
  "mvMORPH" = "32",
  "mvSLOUCH" = "32"
)
method_lw <- c(
  "SMB" = 1.0,
  "Stan" = 0.5,
  "Dense Kalman" = 0.5,
  "Known Schur" = 0.5,
  "PCMBase" = 0.5,
  "PCMFit" = 0.5,
  "mvMORPH" = 0.5,
  "mvSLOUCH" = 0.5
)

bench_panel <- function(data, methods, y_limits) {
  d <- data %>%
    filter(method %in% methods) %>%
    mutate(method = factor(method, levels = methods))
  d_real <- filter(d, !is_extrap)
  d_fake <- filter(d, is_extrap)

  p <- ggplot(
    d_real,
    aes(
      K,
      median_s,
      colour = method,
      shape = method,
      alpha = method,
      linetype = method,
      linewidth = method
    )
  ) +
    geom_line(linewidth = 0.7) +
    geom_point(
      data = d_real,
      aes(x = K, y = median_s),
      inherit.aes = FALSE,
      colour = "white",
      alpha = 1,
      size = 1.9,
      stroke = 0
    ) +
    geom_point(size = 1.9) +
    scale_x_continuous(
      trans = "log2",
      breaks = panel_k,
      labels = as.character(panel_k),
      expand = expansion(mult = c(0.05, 0.14))
    ) +
    scale_y_log10(
      breaks = 10^(-6:6),
      labels = function(x) {
        pow <- log10(x)
        ifelse(
          pow %% 2 == 0,
          scales::math_format(10^.x)(pow),
          ""
        )
      },
      minor_breaks = NULL
    ) +
    coord_cartesian(ylim = y_limits) +
    scale_colour_manual(values = method_cols[methods], name = NULL) +
    scale_shape_manual(values = method_shps[methods], name = NULL) +
    scale_alpha_manual(values = method_alpha[methods], name = NULL) +
    scale_linetype_manual(values = method_lty[methods], name = NULL) +
    scale_linewidth_manual(values = method_lw[methods], name = NULL) +
    labs(x = NULL, y = NULL) +
    theme_classic(base_size = fszT, base_family = fnt) +
    theme(
      aspect.ratio = 1,
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
      panel.grid.major.y = element_line(color = "grey93", linetype = "33", linewidth = 0.4),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_blank(),
      axis.ticks = element_line(linewidth = 0.3),
      axis.text = element_text(size = fsz, family = fnt, color = "black"),
      axis.title = element_text(size = fszT, family = fnt, color = "black"),
      plot.margin = margin(-2, 4, -2, 1, "pt"),
      legend.position = "none"
    )

  if (nrow(d_fake) > 0) {
    extrap_methods <- unique(as.character(d_fake$method))
    d_bridge <- bind_rows(lapply(extrap_methods, function(meth) {
      last_real <- filter(d_real, method == meth) %>% arrange(K) %>% tail(1)
      fake_meth <- filter(d_fake, method == meth) %>% arrange(K)
      bind_rows(last_real, fake_meth)
    }))
    p <- p +
      geom_line(
        data = d_bridge,
        aes(x = K, y = median_s, colour = method, linetype = method, linewidth = method),
        alpha = 0.6,
        inherit.aes = FALSE
      ) +
      geom_point(
        data = d_fake,
        aes(x = K, y = median_s),
        inherit.aes = FALSE,
        colour = "white",
        alpha = 0.6,
        size = 1.9,
        stroke = 0
      ) +
      geom_point(
        data = d_fake,
        aes(x = K, y = median_s, colour = method, shape = method),
        alpha = 0.6,
        size = 1.9,
        inherit.aes = FALSE
      )
  }

  p
}

y_lab <- wrap_elements(
  grid::textGrob(
    "Time (s)",
    rot = 90,
    gp = grid::gpar(fontsize = fszT, fontfamily = fnt)
  )
)

rl <- function(lab) {
  ggplot() +
    annotate(
      "text",
      x = 0.5,
      y = 0.5,
      label = lab,
      angle = 90,
      size = 4,
      fontface = "bold",
      family = fnt,
      hjust = 0.5
    ) +
    theme_void()
}

x_lab <- wrap_elements(
  full = grid::textGrob(
    "State dimension (K)",
    y = grid::unit(1, "npc"),
    just = c("centre", "top"),
    gp = grid::gpar(fontsize = fszT, fontfamily = fnt)
  )
)

top_title_theme <- theme(
  plot.title = element_text(
    hjust = 0.5,
    size = fszT,
    family = fnt,
    face = "bold",
    margin = margin(0, 0, 1, 0, "pt")
  )
)

make_leg_panel <- function(methods, labels = methods) {
  c1x <- 0.12
  c2x <- 0.45
  seg <- 0.035
  gap <- 0.020
  has_proposed <- "SMB" %in% methods
  proposed_methods <- if (has_proposed) "SMB" else character()
  benchmark_methods <- setdiff(methods, proposed_methods)
  proposed_labels <- labels[match(proposed_methods, methods)]
  benchmark_labels <- labels[match(benchmark_methods, methods)]

  leg_df <- tibble(
    method = factor(c(proposed_methods, benchmark_methods), levels = names(method_cols)),
    label = c(proposed_labels, benchmark_labels),
    x_seg0 = c(
      rep(c1x, length(proposed_methods)),
      rep(if (has_proposed) c2x else c1x, length(benchmark_methods))
    ),
    y = c(
      seq(1.4, by = -0.8, length.out = length(proposed_methods)),
      seq(1.8, by = -0.8, length.out = length(benchmark_methods))
    )
  ) %>%
    mutate(
      x_lab = x_seg0 + seg + gap
    )

  if (!has_proposed) {
    leg_df <- leg_df %>% mutate(y = seq(1.8, by = -0.8, length.out = n()))
  }

  leg_titles <- if (has_proposed) {
    tibble(
      label = c("Proposed method", "Benchmarks"),
      x = c(c1x, c2x),
      y = c(2.8, 2.8)
    )
  } else {
    tibble(
      label = "Benchmarks",
      x = c1x,
      y = 2.8
    )
  }

  ggplot() +
    geom_text(
      data = leg_titles,
      aes(x = x, y = y, label = label),
      hjust = 0,
      size = 3.5,
      family = fnt,
      fontface = "bold",
      colour = "grey10",
      inherit.aes = FALSE
    ) +
    geom_segment(
      data = leg_df,
      aes(
        x = x_seg0,
        xend = x_seg0 + seg,
        y = y,
        yend = y,
        colour = method,
        alpha = method,
        linetype = method
      ),
      linewidth = 0.8,
      show.legend = FALSE,
      inherit.aes = FALSE
    ) +
    geom_text(
      data = leg_df,
      aes(x = x_lab, y = y, label = label),
      hjust = 0,
      size = 3.5,
      family = fnt,
      colour = "grey20",
      inherit.aes = FALSE
    ) +
    scale_colour_manual(values = method_cols[methods]) +
    scale_alpha_manual(values = method_alpha[methods]) +
    scale_linetype_manual(values = method_ltyLegend[methods]) +
    scale_x_continuous(limits = c(0, 0.95)) +
    scale_y_continuous(limits = c(-0.9, 3.3)) +
    theme_void() +
    theme(legend.position = "none")
}

make_panel_row <- function(data, row_label, methods, labels = methods, y_limits) {
  panel_list <- lapply(panel_n, function(n_value) {
    bench_panel(filter(data, N == n_value), methods, y_limits = y_limits) +
      labs(title = paste0("N = ", scales::comma(n_value))) +
      top_title_theme
  })

  wrap_plots(
    y_lab,
    rl(row_label),
    panel_list[[1]],
    panel_list[[2]],
    panel_list[[3]],
    plot_spacer(),
    x_lab,
    plot_spacer(),
    make_leg_panel(methods, labels),
    design = "
ABCDE
#FGGG
#HIII
",
    widths = c(0.06, 0.05, 1, 1, 1),
    heights = c(1, 0.05, 0.24)
  )
}

make_two_row_panel <- function(data, row_labels, methods, labels = methods, y_limits) {
  panel_list <- lapply(row_labels, function(row_label) {
    lapply(panel_n, function(n_value) {
      bench_panel(
        filter(data, operation == row_label, N == n_value),
        methods,
        y_limits = y_limits
      ) +
        labs(title = paste0("N = ", scales::comma(n_value))) +
        top_title_theme
    })
  })

  wrap_plots(
    y_lab,
    rl(row_labels[[1L]]),
    panel_list[[1L]][[1L]],
    panel_list[[1L]][[2L]],
    panel_list[[1L]][[3L]],
    rl(row_labels[[2L]]),
    panel_list[[2L]][[1L]],
    panel_list[[2L]][[2L]],
    panel_list[[2L]][[3L]],
    plot_spacer(),
    x_lab,
    plot_spacer(),
    make_leg_panel(methods, labels),
    design = "
ABCDE
AFGHI
#JKKK
#LMMM
",
    widths = c(0.06, 0.05, 1, 1, 1),
    heights = c(1, 1, 0.05, 0.24)
  )
}

panel_y_limits <- function(data, lower = 2.5e-4, pad = 1.8) {
  values <- data$median_s
  values <- values[is.finite(values) & values > 0]
  if (length(values) == 0L) {
    return(c(lower, 1))
  }
  c(min(lower, min(values) / pad), max(values) * pad)
}
```

# Time-Series Benchmarks

```{r time-series-figure, eval=FALSE}
time_series_methods <- c("SMB", "Stan", "Dense Kalman", "Known Schur")
time_series_methods <- time_series_methods[
  time_series_methods %in% as.character(unique(time_series_timings$method))
]
if (length(time_series_methods) == 0L) {
  stop("No time-series benchmark rows were found.", call. = FALSE)
}

time_series_fig <- make_two_row_panel(
  time_series_timings,
  row_labels = c("Likelihood", "Gradient"),
  methods = time_series_methods,
  labels = time_series_methods,
  y_limits = panel_y_limits(time_series_timings)
)

time_series_fig
```

```{r time-series-save, eval=FALSE}
save_benchmark_pdf(time_series_fig, "ou_time_series_benchmarks", width = 10, height = 8)
```

# Tree Benchmarks

```{r tree-figure, eval=FALSE}
tree_methods <- c("SMB", "PCMBase", "PCMFit", "mvMORPH", "mvSLOUCH")
tree_methods <- tree_methods[
  tree_methods %in% as.character(unique(tree_timings$method))
]
if (length(tree_methods) == 0L) {
  stop("No tree benchmark rows were found.", call. = FALSE)
}

tree_fig <- make_panel_row(
  tree_timings,
  row_label = "Trees",
  methods = tree_methods,
  labels = tree_methods,
  y_limits = panel_y_limits(tree_timings)
)

tree_fig
```

```{r tree-save, eval=FALSE}
save_benchmark_pdf(tree_fig, "ou_tree_benchmarks", width = 10, height = 4.25)
```

# Output

Figures are written to:

- `r figure_output_paths`
